---
title: "1. Exploratory data analysis"
author: "Inigo Bermejo"
date: "2024-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploring the NHEFS dataset

## Background

The NHEFS dataset is used in the Hernan and Robins book [1]. The acronym that stands for National Health and Nutrition Examination Survey Data I Epidemiologic Follow-up Study. We will use this dataset to estimate the effect of smoking cessation on weight gain.

The codebook of the dataset can be found [here](https://www.hsph.harvard.edu/miguel-hernan/wp-content/uploads/sites/1268/2012/10/NHEFS_Codebook.xls). A detailed description of the NHEFS, can be found at <http://wwwn.cdc.gov/nchs/nhanes/nhefs/>.

## Loading the dataset into R

Let us load the dataset into R using the read.csv function.

```{r}
nhefs <- read.csv("..\\data\\NHEFS.csv")
```

## Exploring the variables

Let us print the available variables:

```{r}
colnames(nhefs)
```

And focus on the important ones

-   **qsmk**: Quit smoking between 1st questionnaire and 1982, 1:yes, 0:no
-   **age**: age in 1971
-   **sex**: 0: male 1: female
-   **smokeintensity**: Increase in number of cigarettes/day between 1971 and 1982
-   **smokeyrs**: Years of smoking
-   **wt71**: Weight in kilograms in 1971
-   **wt82**: Weight in kilograms in 1982

## Summary statistics

First thing, we will print the summary statistics of the key variables:

```{r}
# provisionally ignore subjects with missing values for weight in 1982
nhefs.nmv <- nhefs[which(!is.na(nhefs$wt82)),]

# Summary
cat("Age:\n")
summary(nhefs.nmv$age)
cat("Gender distribution (%):")
round(prop.table(table(nhefs.nmv$sex))*100,1)
cat("Weight on 71:\n")
summary(nhefs.nmv$wt71)
cat("Increase in number of cigarrettes / day:\n")
summary(nhefs.nmv$smokeintensity)
cat("Years of smoking:\n")
summary(nhefs.nmv$smokeyrs)

### MISSING CODE HERE
```

## Bivariate analyses

Now let us look at the interplay between some variable pairs:

```{r}
cat("qsmk vs sex:")
table(nhefs.nmv$qsmk, nhefs.nmv$sex)
round(prop.table(table(nhefs.nmv$qsmk, nhefs.nmv$sex), 1), 2)

### MISSING CODE HERE

```

## References

1.  Hernán MA, Robins JM (2020). Causal Inference: What If. Boca Raton: Chapman & Hall/CRC.
