---
title: "3. Propensity score matching"
author: "Inigo Bermejo"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3. Propensity score matching

## Setup

Install and load required packages:

```{r}
library(tableone)
library(Matching)
library(MatchIt)

```

## Predictor imbalance

```{r}
xvars <- c("sex", "race", "age", "education", "smokeintensity", "smokeyrs", "exercise", "active")

table1 <- CreateTableOne(vars=xvars, data= nhefs.nmv, strata="qsmk", test = FALSE)

print(table1, smd=TRUE)

```

## Propensity score model

Fit the propensity score model

```{r}
psmodel <- glm(
  ### MISSING CODE HERE
  family = binomial(),
  data = nhefs.nmv)

summary(psmodel)

```

## Assign propensity scores to data

```{r}

pscore <- psmodel$fitted.values

summary(pscore)

```

## Match samples based on propensity score

```{r}

psmatch <- Match(Tr=nhefs.nmv$qsmk, M=1, X=pscore, replace = FALSE)

matched <- nhefs.nmv[unlist(psmatch[c("index.treated", "index.control")]),]

cat(paste0('Matched pairs:', nrow(matched), '\n'))
```

Check the matched table:

```{r}
matchedtab1 <- CreateTableOne(vars= xvars, strata="qsmk", data=matched, test = FALSE)

print(matchedtab1, smd=TRUE)

```

## Use a caliper 

Sometimes we do not want to force all the samples to be paired. In those cases, we use a caliper to set the maximum distance that is acceptable for each match.

```{r}
psmatch <- Match(Tr=nhefs.nmv$qsmk, M=1, X=pscore, replace = FALSE, caliper = 0.1)

matched <- nhefs.nmv[unlist(psmatch[c("index.treated", "index.control")]),]

cat(paste0('Matched pairs:', nrow(matched), '\n'))

matchedtab1 <- CreateTableOne(vars= xvars, strata="qsmk", data=matched, test = FALSE)

print(matchedtab1, smd=TRUE)

```

## Estimate ATE

```{r}
### MISSING CODE HERE
```

## Using PS in regression as coefficient

```{r}
summary(glm(wt82_71 ~ qsmk ### MISSING CODE HERE
              , data = nhefs.nmv))
```
