---
title: "5. Standardisation and the parametric G-formula"
author: "Inigo Bermejo"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5. Standardisation and the parametric G-formula

## Standardizing the mean outcome to the baseline confounders

Create a dataset with 3 copies of each subject

```{r}

nhefs$interv <- -1 # 1st copy: equal to original one

interv0 <- nhefs # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0
interv0$qsmk <- 0
interv0$wt82_71 <- NA

interv1 <- nhefs # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1
interv1$qsmk <- 1
interv1$wt82_71 <- NA

onesample <- rbind(nhefs, interv0, interv1) # combining datasets
```

Fit a linear model to estimate mean outcome conditional on treatment and confounders. Parameters are estimated using original observations only (nhefs)

```{r}

std <- glm(
  ### MISSING CODE HERE
  data = onesample
)
summary(std)
```

Use the model to predict outcome for observations with treatment set to 0 (interv=0) and to 1 (interv=1)

```{r}
onesample$predicted_Y <- ### MISSING CODE HERE
```

Estimate mean outcome in each of the groups interv=0, and interv=1 This mean outcome is a weighted average of the mean outcomes in each combination of values of treatment and confounders, that is, the standardized outcome

```{r}
original     <- mean(onesample[which(onesample$interv == -1), ]$predicted_Y)

none_treated <- mean(onesample[which(onesample$interv == 0), ]$predicted_Y)

all_treated  <- mean(onesample[which(onesample$interv == 1), ]$predicted_Y)

cat(paste0("Predicted weight gain in original dataset:     " , round(original, 4), "\n"))
cat(paste0("Predicted weight gain if no one quits smoking: " , round(none_treated, 4), "\n"))
cat(paste0("Predicted weight gain if all quit smoking:     " , round(all_treated, 4), "\n"))
cat(paste0("Estimated ATE:                                 " , round(all_treated-none_treated, 4), "\n"))

```

## Computing the 95% confidence interval of the standardized means and their difference

```{r}
library(boot)

# function to calculate difference in means
standardisation <- function(data, indices) {
  # create a dataset with 3 copies of each subject
  d <- data[indices, ] # 1st copy: equal to original one`
  d$interv <- -1
  d0 <- d # 2nd copy: treatment set to 0, outcome to missing
  d0$interv <- 0
  d0$qsmk <- 0
  d0$wt82_71 <- NA
  d1 <- d # 3rd copy: treatment set to 1, outcome to missing
  d1$interv <- 1
  d1$qsmk <- 1
  d1$wt82_71 <- NA
  d.onesample <- rbind(d, d0, d1) # combining datasets

  # linear model to estimate mean outcome conditional on treatment and confounders
  # parameters are estimated using original observations only (interv= -1)
  # parameter estimates are used to predict mean outcome for observations with set
  # treatment (interv=0 and interv=1)
  fit <- glm(
  ### MISSING CODE HERE
    data = d.onesample
  )

  d.onesample$predicted_Y <- ### MISSING CODE HERE

  # estimate mean outcome in each of the groups interv=-1, interv=0, and interv=1
  return(c(
    mean(d.onesample$predicted_Y[d.onesample$interv == -1]),
    mean(d.onesample$predicted_Y[d.onesample$interv == 0]),
    mean(d.onesample$predicted_Y[d.onesample$interv == 1]),
    mean(d.onesample$predicted_Y[d.onesample$interv == 1]) -
      mean(d.onesample$predicted_Y[d.onesample$interv == 0])
  ))
}

# bootstrap
results <- boot(data = nhefs,
                statistic = standardisation,
                R = 5)

# generating confidence intervals
se <- c(sd(results$t[, 1]),
        sd(results$t[, 2]),
        sd(results$t[, 3]),
        sd(results$t[, 4]))
mean <- results$t0
ll <- mean - qnorm(0.975) * se
ul <- mean + qnorm(0.975) * se

bootstrap <-
  data.frame(cbind(
    c(
      "Observed",
      "No Treatment",
      "Treatment",
      "Treatment - No Treatment"
    ),
    mean,
    se,
    ll,
    ul
  ))
bootstrap
```
