---
title: "2. Calculating ATE (naive)"
author: "Inigo Bermejo"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Calculating ATE (naive)

We will start estimating ATE following a naive approach, assuming no confounding.

We will calculate the difference between those who stopped smoking and those who did not.

First, we calculate the average change in weight for the total population

```{r}

mean(nhefs.nmv$wt82_71)

```

Then, we will calculate the average change in those who stopped smoking and those who did not.

```{r}

wt82_71_qsmk_0 <- mean(nhefs.nmv$wt82_71[nhefs.nmv$qsmk == 0])

wt82_71_qsmk_1 <- mean(nhefs.nmv$wt82_71[nhefs.nmv$qsmk == 1])

cat(paste0('Weight gain (quit smoking = no): ', 
           round(wt82_71_qsmk_0, 4),'\n', 
           'Weight gain (quit smoking = yes): ', 
           round(wt82_71_qsmk_1, 4),'\n'))

```

The naive ATE is the difference between the two:

```{r}
cat(paste0("Naive ATE estimation: ", round(wt82_71_qsmk_1 - wt82_71_qsmk_0, 3), '\n'))
```

We can reach the same result fitting a linear model:

```{r}

summary(glm(wt82_71 ~ qsmk, data = nhefs.nmv))
```
