---
title: "4. IPW and MSM"
author: "Inigo Bermejo"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Inverse probability weighting (IPW) and Marginal Structural Models (MSM)

## Estimation of IP weights via a logistic model

First, fit the model to estimate the probability of treatment:

```{r}
fit <- glm(
  qsmk ~ sex + race + age + I(age ^ 2) +
    as.factor(education) + smokeintensity +
    I(smokeintensity ^ 2) + smokeyrs + I(smokeyrs ^ 2) +
    as.factor(exercise) + as.factor(active) + wt71 + I(wt71 ^ 2),
  family = binomial(),
  data = nhefs.nmv
)
summary(fit)
```

Next, calculate the weights.

First, assign probability of being treated or not being treated, depending on the observed treatment allocation

```{r}
p.qsmk.obs <-
  ifelse(nhefs.nmv$qsmk == 0,
         1 - predict(fit, type = "response"),
         predict(fit, type = "response"))
```

Finally, calculate the inverse of the probability:

```{r}
nhefs.nmv$w <- 1 / p.qsmk.obs
summary(nhefs.nmv$w)
sd(nhefs.nmv$w)
```

## Apply the weights and fit the marginal structural model

Using the 'geepack' package

```{r}
# install.packages("geepack") # install package if required
library("geepack")
msm.w <- geeglm(
  wt82_71 ~ qsmk,
  data = nhefs.nmv,
  weights = w,
  id = seqn,
  corstr = "independence"
)
summary(msm.w)

```

Calculate confidence interval

```{r}
beta <- coef(msm.w)
SE <- coef(summary(msm.w))[, 2]
lcl <- beta - qnorm(0.975) * SE
ucl <- beta + qnorm(0.975) * SE
cbind(beta, lcl, ucl)

```

Using the 'survey' package

```{r}

library(survey)

weighteddata <- svydesign(ids= ~ 1, data=nhefs.nmv, weights = ~ w)

msm <- (svyglm(wt82_71 ~ qsmk, design= weighteddata))

coef(msm)
confint(msm)
```
