---
title: "5. Standardisation and the parametric G-formula"
author: "Inigo Bermejo"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5. Standardisation and the parametric G-formula

## Estimating the mean outcome within levels of treatment and confounders:

Fit a model to predict the outcome, using treatment and confounders:

```{r}
# some preprocessing of the data
nhefs$cens <- ifelse(is.na(nhefs$wt82), 1, 0)

fit <-
  glm(
    wt82_71 ~ qsmk + sex + race + age + I(age * age) + as.factor(education)
    + smokeintensity + I(smokeintensity * smokeintensity) + smokeyrs
    + I(smokeyrs * smokeyrs) + as.factor(exercise) + as.factor(active)
    + wt71 + I(wt71 * wt71) + qsmk * smokeintensity,
    data = nhefs
  )
summary(fit)
```


## Standardizing the mean outcome to the baseline confounders

Create a dataset with 3 copies of each subject

```{r}

nhefs$interv <- -1 # 1st copy: equal to original one

interv0 <- nhefs # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0
interv0$qsmk <- 0
interv0$wt82_71 <- NA

interv1 <- nhefs # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1
interv1$qsmk <- 1
interv1$wt82_71 <- NA

onesample <- rbind(nhefs, interv0, interv1) # combining datasets

# linear model to estimate mean outcome conditional on treatment and confounders
# parameters are estimated using original observations only (nhefs)
# parameter estimates are used to predict mean outcome for observations with
# treatment set to 0 (interv=0) and to 1 (interv=1)

std <- glm(
  wt82_71 ~ qsmk + sex + race + age + I(age * age)
  + as.factor(education) + smokeintensity
  + I(smokeintensity * smokeintensity) + smokeyrs
  + I(smokeyrs * smokeyrs) + as.factor(exercise)
  + as.factor(active) + wt71 + I(wt71 * wt71) + I(qsmk * smokeintensity),
  data = onesample
)
summary(std)
```



Estimate mean outcome in each of the groups interv=0, and interv=1
This mean outcome is a weighted average of the mean outcomes in each combination
of values of treatment and confounders, that is, the standardized outcome
```{r}
original     <- mean(onesample[which(onesample$interv == -1), ]$predicted_meanY)

none_treated <- mean(onesample[which(onesample$interv == 0), ]$predicted_meanY)

all_treated  <- mean(onesample[which(onesample$interv == 1), ]$predicted_meanY)

cat(paste0("Predicted weight gain in original dataset:     " , round(original, 4), "\n"))
cat(paste0("Predicted weight gain if no one quits smoking: " , round(none_treated, 4), "\n"))
cat(paste0("Predicted weight gain if all quit smoking:     " , round(all_treated, 4), "\n"))
cat(paste0("Estimated ATE:                                 " , round(all_treated-none_treated, 4), "\n"))

```

## Computing the 95% confidence interval of the standardized means and their difference

```{r}
library(boot)

# function to calculate difference in means
standardisation <- function(data, indices) {
  # create a dataset with 3 copies of each subject
  d <- data[indices, ] # 1st copy: equal to original one`
  d$interv <- -1
  d0 <- d # 2nd copy: treatment set to 0, outcome to missing
  d0$interv <- 0
  d0$qsmk <- 0
  d0$wt82_71 <- NA
  d1 <- d # 3rd copy: treatment set to 1, outcome to missing
  d1$interv <- 1
  d1$qsmk <- 1
  d1$wt82_71 <- NA
  d.onesample <- rbind(d, d0, d1) # combining datasets

  # linear model to estimate mean outcome conditional on treatment and confounders
  # parameters are estimated using original observations only (interv= -1)
  # parameter estimates are used to predict mean outcome for observations with set
  # treatment (interv=0 and interv=1)
  fit <- glm(
  wt82_71 ~ qsmk + sex + race + age + I(age * age)
    + as.factor(education) + smokeintensity
    + I(smokeintensity * smokeintensity) + smokeyrs
    + I(smokeyrs * smokeyrs) + as.factor(exercise)
    + as.factor(active) + wt71 + I(wt71 * wt71) + I(qsmk * smokeintensity),
    data = d.onesample
  )

  d.onesample$predicted_meanY <- predict(fit, d.onesample)

  # estimate mean outcome in each of the groups interv=-1, interv=0, and interv=1
  return(c(
    mean(d.onesample$predicted_meanY[d.onesample$interv == -1]),
    mean(d.onesample$predicted_meanY[d.onesample$interv == 0]),
    mean(d.onesample$predicted_meanY[d.onesample$interv == 1]),
    mean(d.onesample$predicted_meanY[d.onesample$interv == 1]) -
      mean(d.onesample$predicted_meanY[d.onesample$interv == 0])
  ))
}

# bootstrap
results <- boot(data = nhefs,
                statistic = standardisation,
                R = 5)

# generating confidence intervals
se <- c(sd(results$t[, 1]),
        sd(results$t[, 2]),
        sd(results$t[, 3]),
        sd(results$t[, 4]))
mean <- results$t0
ll <- mean - qnorm(0.975) * se
ul <- mean + qnorm(0.975) * se

bootstrap <-
  data.frame(cbind(
    c(
      "Observed",
      "No Treatment",
      "Treatment",
      "Treatment - No Treatment"
    ),
    mean,
    se,
    ll,
    ul
  ))
bootstrap
```
