---
title: "6. Doubly robust estimator"
author: "Inigo Bermejo"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 6. Doubly robust estimator

We have seen that the doubly robust (DR) estimator is unbiased for the ATE if at least one of the models (i.e. the outcome model or the propensity score model) are correctly specified. Moreover If both models are correctly specified, the DR estimator is semiparametrically efficient.

## Fit the propensity score model

```{r}
psmodel <- glm(
  qsmk ~ sex + race + age + I(age ^ 2) +
    as.factor(education) + smokeintensity +
    I(smokeintensity ^ 2) + smokeyrs + I(smokeyrs ^ 2) +
    as.factor(exercise) + as.factor(active) + wt71 + I(wt71 ^ 2),
  family = binomial(),
  data = nhefs.nmv)

summary(psmodel)
```

## Fit the outcome model

```{r}
nhefs$cens <- ifelse(is.na(nhefs$wt82), 1, 0)

outcome_model <-
  glm(
    wt82_71 ~ qsmk + sex + race + age + I(age * age) + as.factor(education)
    + smokeintensity + I(smokeintensity * smokeintensity) + smokeyrs
    + I(smokeyrs * smokeyrs) + as.factor(exercise) + as.factor(active)
    + wt71 + I(wt71 * wt71) + qsmk * smokeintensity,
    data = nhefs
  )
summary(outcome_model)

```

## Doubly robust estimation

Data preparation

```{r}

# Propensity scores
pi.hat <- predict(psmodel, type = 'response', newdata = nhefs.nmv)

# Outcome model predictions
nhefs.0 <- nhefs.nmv
nhefs.0$qsmk <- 0
nhefs.1 <- nhefs.nmv
nhefs.1$qsmk <- 1
b.hat.0<- predict(outcome_model,newdata = nhefs.0)
b.hat.1<- predict(outcome_model,newdata = nhefs.1)

# Renaming variables for convenience
a<-nhefs.nmv$qsmk
y<-nhefs.nmv$wt82_71
```

Estimation based on formulas in slide 164.

```{r}
e.hat.1 <-(a*y/pi.hat - (a/pi.hat-1)*b.hat.1)
e.hat.0 <- (1-a)*y/(1-pi.hat) - ((1-a)/(1-pi.hat)-1)*b.hat.0
e.hat <-e.hat.1-e.hat.0
beta.hat<-mean(e.hat)

var.beta.hat<-var(e.hat)/nrow(nhefs)

cat(paste0("ATE estimate:", round(beta.hat, 4), '\n'))

```
